---
title: "ceLLama"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
ceLLama is a simple automation pipeline for cell type annotations using large-language models (LLMs). 

It has several advantages:

- Works locally, thus no information leak.
- Takes negative genes into account. 
- Quite fast.
- Trained on almost all internet!

ceLLama can be super useful for quick & dirty cell type checks!

## How to use

First, you need to download [`ollama`](https://ollama.com/).

Then you can choose the model of your choice. 
Currently, one of the best open source LLM models is Llama3. 
You can run it on your terminal simply using:

```{bash eval=FALSE}
ollama run llama3
```

This starts a local server on your machine, and you can see if it is running by checking http://localhost:11434/. It should say "Ollama is running".

Then you are ready to go!

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(httr)

pbmc.data <- Read10X("../../Downloads/filtered_gene_bc_matrices/hg19/")

pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)

pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")

pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)

# note that you can chain multiple commands together with %>%
pbmc <- SCTransform(pbmc, verbose = F) %>%
    RunPCA(verbose = F) %>%
    FindNeighbors(dims = 1:10, verbose = F) %>%
    FindClusters(resolution = 0.5, verbose = F) %>% 
    RunUMAP(dims = 1:10, verbose = F)

DimPlot(pbmc, label = T, label.size = 3) + theme_linedraw() + theme(aspect.ratio = 1)
```


```{r find DEGs}
# Find cluster markers
pbmc.markers <- FindAllMarkers(pbmc, verbose = F)

# split into a lists per cluster
pbmc.markers.list <- split(pbmc.markers, pbmc.markers$cluster)
```

```{r}
# run cellama!
# set seed, make temperature 0 for reproducible results
library(ceLLama)

res <- ceLLama(pbmc.markers.list, temperature = 0, seed = 101)
```

```{r}
# transfer the labels
annotations <- map_chr(res, 1)

names(annotations) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, annotations)

DimPlot(pbmc, label = T, repel = T, label.size = 3) + theme_linedraw() + theme(aspect.ratio = 1)
```

```{r eval=FALSE}
# Get the reason for the annotation! (a bit slower)
res <- ceLLama(pbmc.markers.list, temperature = 0, seed = 101, get_reason = T)
```


